Master thesis

### 2017-06-08 docs/master-thesis.tex
Model and Appendix

Carbon assimilation is computed over half-hourly period of a
representative day. Then allocation is computed to simulate tree growth
from an explicit carbone balance (in contrast to previous models).
Finally environment is updated at each timestep set to one month.
Seedlings are not simulated explicitly but as a pool. In addition
belowground processes, herbaceous plants, epiphytes and lianas are not
simulated inside TROLL. The source code is written in C++ and available
upon request. All modules of TROLL models are further detailed in
\protect\hyperlink{appendix-1}{Appendix 1}.

@@

Previous implementation of TROLL model used \citet{Reich1991a} allometry
to infer leaf lifespan \(LL\) from species leaf mass per area \(LMA\)
\citep[see \protect\hyperlink{appendix-1}{Appendix 1}]{Li}. But the use
of the allometrie from \citet{Reich1991a} with current implementation of
the TROLL model resulted in an underestimation of leaf lifespan for low
LMA species. Consequently in the following paragraph we suggest a new
allometry.

@@

Leaf lifespan \(LL\) follows a lognormal law with location infered from
leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
presemption following a gamma law. \(d\) represents the dataset random
effects and encompass environmental and protocol variations. \textbf{Add
Appendix ?}

@@

\textbf{Add Appendix ?}

@@

Due to crown aspects, treefall from logs are often random (whereas
difficult to manage, treefall can still be oriented, Laurent Descroix,
ONF, personnal communication). Consequently, TROLL consider treefall
from log as random like current natural treefall implementation inside
TROLL (see code \protect\hyperlink{appendix-1}{Appendix 1}).

@@

\(Death\) of a tree \(i\) follows a \(\mathcal{Bernoulli}\) law of
probability \(P(Death_i)\). The odds for a tree \(i\) to die
\(O(Death_i)\) is calculated by the sum of the natural tree death odd
\(\theta\) and a perturbation index \(\beta*e^{\alpha*d_{gaps}}\). The
perturbation index depend on the distance \(d_{gaps}\) of the tree \(i\)
to the closest logging gap. The probability for a tree \(i\) to die
\(P(Death_i)\) is finally calculated by taking the inverse logit of the
odd \(logit^{-1}[O(Death_i)]\). \textbf{Add Appendix ?}

@@

\citet{Li} already assessed TROLL model sensitivity to several
parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
key role in model functioning.

On the other hand, we decided to use TROLL to study resistance and
resilience of ecosystem face to disturbance, highlighting the role of
biodiversit. Consequently we particulrarly needed to assess the
importance of functional traits to further better control and evaluate
functional diversities. We also needed to assess the sensitivity of
TROLL model to the seed rain constant (\(n_{ext}\), see \eqref{eq:next})
because we assume it is one of the main factor of tree recruitments
after disturbance in simulations.

TROLL model currenty uses leaf mass per area (\(LMA\) in \(g.m^{-2}\)),
leaf nitrogen content per dry mass (\(N_m\) in \(mg.g^{-1}\)), leaf
phosphorus content per dry mass (\(P_m\) in \(mg.g^{-1}\)), wood
specific gravity (\(wsg\) in \(g.cm^{-3}\)), diameter at breasth height
threshold (\(dbh_{thresh}\) in \(m\)), asymptotic height (\(h_{lim}\) in
\(m\)), and parameter of the tree-height-dbh allometry (\(a_h\) in
\(m\)). To assess the sensitivity of TROLL model to species functionnal
traits, we performed a sensitivity analysis by fixing species trait
values to their mean. Each trait was tested independently. We reduce to
a common mean traits with a Perason's correlation value \(r \geq 0.8\)
(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)).

To assess the sensitivity of TROLL model to seed rain, we performed a
sensitivity analysis by fixing simulations seed rain constant to 2, 20,
200 and 2000 seeds per hectare (default value being 200 seeds per
hectare).

@@

\emph{Most tropical timber production originates from selective logging,
the targeted harvesting of timber from species of interest.
Consequently, tropical sylviculture can be assimilated to a disturbance.
The main difference between a disturbance and selective logging is the
targetting of both species and individuals of interest. So we decided to
first asses unselective disturbance effect on tropical forest ecosystem
to subsequently better understand selective logging effect. We first
implemented a simple disturbance module inside TROLL model to simulate
unselective disturbance.} \textbf{Already inside model descripiton !}

We first implemented a disturbance module inside TROLL model to simulate
unselective disturbance. 60 tree communities were then defined with
different levels of both specific and functional richness to explore the
role of biodiversity on tropical forest ecosystem answer to disturbance.
Next, different levels of disturbance were simulated on previous mature
communities. Finally, resilience of ecosystem major global variables
(related to carbon stock, forest dynamic and florisitic structure) from
simulations were partitioned between selection effects and niche
complementarity \citep{Loreau2001}.

\subsection{Selective logging}\label{selective-logging}

Once the unselective effect of disturbance on forest ecosystem studied,
we can focus on selective logging to highlight the effect of species and
individuals targetting on the ecosystem answer. We first implemented a
sylviculture module inside TROLL model to simulate selective logging in
regards to french Guiana practices. We used the 60 communities from
disturbance simulations to study the role of biodiversity in both forest
ecosystem and commercial species answer to selective logging. We then

@@

In order to assess the role of biodiversity in ecosystem answer to
disturbance, we needed to create a space of experiments encompassing
both variation of disturbance, biodiversity and time. Disturbance was
represented by percentage of basal area loss (0\%, 25\%, 50\% and 75\%).
Biodiversity was integrated with two components specific and functional
diversities. We used species richness \(SR\) to represents species
diversity (5, 25, and 125). Functional diversity can be related to
numerous components, and \citet{Borgy2017} argued for 5: richness,
divergence, regularity, overlap and mean. Because mature forest were
created from a bare soil with TROLL simulations, we could not control a
priori divergence, regularity and overlap but only assess theim before
diturbance. Consequently, we focused on functional richness with convex
hull volume \(CHV\) and functional mean with community weighted mean
\(CWM\). For each level of species richness \(SR\), we selected 20
communities with growing convex hull volume \(CHV\) but with a community
weighted means close to the regional species pool community weighted
means. Effectivelly, we did not wanted drastic change in community means
that could have more effect than functional richness itself. This design
of experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
(maturity being assumed after 500 years of regeneration \citep{Li}).
Figure \ref{fig:DOE} presents the design of experiment for communities
biodiversity after the forest mature were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/DOE-1.pdf}
\caption{\label{fig:DOE}Experimental design before disturbance. Communities
are implemented along a gradient of species richness (SR) and functional
dispersion (FDis) resulting in a broad range of aboveground biomass
(AGB). FDis was caluclated based on 4 functional traits (leaf mass per
area, wood specific gravity, maximum diameter, maximum height).}
\end{figure}

\subsection{Ecosystem answer anlaysis}\label{ecosystem-answer-anlaysis}

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We focused on few functions and related
metrics to analyse ecosystem answer to disturbance: carbon stock with
aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), forest dynamic with
number of stem above \(10~cm\) diameter at breast height (\(N10\)), and
floristic composition with \textbf{?} (\textbf{?} in \textbf{?}).

@@

\begin{equation}
  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
  \label{eq:Resilience}
\end{equation}

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). We transformed the equation
in the resilience of the system \(R(t)\) at the time \(t\) being the
ratio of the ecosystem metric \(X(t)\) at time \(t\) to the ecosystem
metric before the disturbance happened \(X(t_d-1)\) at time \(t_d-1\).
\(X(t_d-1)\) was calculated as the mean of the ecosystem output for 50
last years of the simulation of the mature forest (over 600 years).
Then, we calculated for each simulation the recovery time
\(t_{recovery}\) were the ecosystem reached back its stable state
defined as its state before disturbance. But some ecosystem metrics can
reach pre disturbance value wihtout revealing ecosystem recovery. For
instance, number of stem above \(10~cm\) diameter at breast height
\(N10\) will first decrease due to disturbance. Then \(N10\) will exceed
its initial value due to new seedlings recruitment before decreasing
towards its pre disturbanec value. Consequently, the first time \(N10\)
reaches back its pre disturbance value can not be considered as the
recovery time \(t_{recovery}\). In order to solve this issue, we
considered the ecosystem as stable again when its whole set of observed
metrics reach their pre disturbance values.

Biodiversity is not only a facet of the experimental design and an
ecosystem output through florisitic composition, but also interact on
ecosystem functioning and consequently on its answer to disturbance.
Biodiversity ecosystem functioning relation can be split in
complementarity and selection effect with \citet{Loreau2001}
partitioning:

@@

\subsection{Disturbance}\label{disturbance-2}

@@

\hypertarget{appendix-1}{\section{Appendix 1}\label{appendix-1}}

@@

\subsection{Leaf lifespan}\label{leaf-lifespan-1}

\begin{equation}
  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
  \label{eq:LLb}
\end{equation}

Leaf lifespan \(LL\) follows a lognormal law with location infered from
leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
presemption following a gamma law. \(d\) represents the dataset random
effects and encompass environmental and protocol variations. The
sampling of model \eqref{eq:LLb} resulted in the following allometry:

\begin{equation}
  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
  \label{eq:LL}
\end{equation}

@@

>>>>

Carbon assimilation is computed over half-hourly period of a
representative day. Then allocation is computed to simulate tree growth
from an explicit carbone balance (in contrast to previous models).
Finally environment is updated at each timestep set to one month.
Seedlings are not simulated explicitly but as a pool. In addition
belowground processes, herbaceous plants, epiphytes and lianas are not
simulated inside TROLL. The source code is written in C++ and available
upon request. All modules of TROLL models are further detailed in
\protect\hyperlink{appendix-1-troll-model}{Appendix 1: TROLL model}.

@@

Previous implementation of TROLL model used \citet{Reich1991a} allometry
to infer leaf lifespan \(LL\) from species leaf mass per area \(LMA\)
\citep[see \protect\hyperlink{appendix-1-troll-model}{Appendix 1: TROLL
model}]{Li}. But the use of the allometrie from \citet{Reich1991a} with
current implementation of the TROLL model resulted in an underestimation
of leaf lifespan for low LMA species. Consequently in the following
paragraph we suggest a new allometry.

@@

Leaf lifespan \(LL\) follows a lognormal law with location infered from
leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
presemption following a gamma law. \(d\) represents the dataset random
effects and encompass environmental and protocol variations (see
\protect\hyperlink{appendix-2-leaf-lifespan-model}{Appendix 2: Leaf
lifespan model} for more details).

@@

(see \protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten
tree model} for more details)

@@

Due to crown aspects, treefall from logs are often random (whereas
difficult to manage, treefall can still be oriented, Laurent Descroix,
ONF, personnal communication). Consequently, TROLL consider treefall
from log as random like current natural treefall implementation inside
TROLL (see code {[}Appendix 1{]}).

@@

\(Death\) of a tree \(i\) follows a \(\mathcal{Bernoulli}\) law of
probability \(P(Death_i)\). The odds for a tree \(i\) to die
\(O(Death_i)\) is calculated by the sum of the natural tree death odd
\(\theta\) and a perturbation index \(\beta*e^{\alpha*d_{gaps}}\). The
perturbation index depend on the distance \(d_{gaps}\) of the tree \(i\)
to the closest logging gap. The probability for a tree \(i\) to die
\(P(Death_i)\) is finally calculated by taking the inverse logit of the
odd \(logit^{-1}[O(Death_i)]\) (see
\protect\hyperlink{appendix-4-gap-damages-model}{Appendix 4: Gap damages
model} for more details).

@@

\citet{Li} already assessed TROLL model sensitivity to several
parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
key role in model functioning. On the other hand, we decided to use
TROLL to study resistance and resilience of ecosystem face to
disturbance, highlighting the role of biodiversity. Consequently we
particulrarly needed to assess the importance of functional traits to
further better control and evaluate functional diversities. We also
needed to assess the sensitivity of TROLL model to the seed rain
constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it as one
of the main factors of tree recruitments after disturbance within
simulations.

TROLL model currenty uses leaf mass per area (\(LMA\) in \(g.m^{-2}\)),
leaf nitrogen content per dry mass (\(N_m\) in \(mg.g^{-1}\)), leaf
phosphorus content per dry mass (\(P_m\) in \(mg.g^{-1}\)), wood
specific gravity (\(wsg\) in \(g.cm^{-3}\)), diameter at breasth height
threshold (\(dbh_{thresh}\) in \(m\)), asymptotic height (\(h_{lim}\) in
\(m\)), and parameter of the tree-height-dbh allometry (\(a_h\) in
\(m\)). To assess the sensitivity of TROLL model to species functionnal
traits, we performed a sensitivity analysis by fixing species trait
values to their mean. Each trait was tested independently. We reduce to
a common mean traits with a Pearson's correlation value \(r \geq 0.8\)
(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)). To assess
the sensitivity of TROLL model to seed rain, we performed a sensitivity
analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000
seeds per hectare.

@@

@@

In order to assess the role of biodiversity in ecosystem answer to both
disturbance and sylviculture, we needed to create a space of experiments
encompassing both variation of disturbance, biodiversity and time.
Disturbance was represented by percentage of basal area loss (0\%, 25\%,
50\% and 75\%), or as a selective logging simulation. Biodiversity was
integrated with two components of its components: taxonomic and
functional diversities. We used species richness \(SR\) to represents
taxonomic diversity (5, 25, and 125 species). Functional diversity can
be related to numerous components, and \citet{Borgy2017} argued for 5:
richness, divergence, regularity, overlap and mean. Because mature
forest were created from a bare soil with TROLL simulations, we could
not control a priori divergence, regularity and overlap but only assess
theim before diturbance. Consequently, we focused on functional richness
with convex hull volume \(CHV\) and functional mean with community
weighted mean \(CWM\). For each level of species richness \(SR\), we
selected 20 communities with growing convex hull volume \(CHV\) but with
a community weighted means close to the regional species pool community
weighted means. Effectivelly, we did not wanted drastic change in
community means that could have more effect than functional richness
itself. This design of experiments resulted in 60 communities
(\(5~SR*20~CHV\)) and 240 simulations
(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
being assumed after 500 years of regeneration \citep{Li}). Figure
\ref{fig:DOE} presents the design of experiment for communities
biodiversity after the mature forest were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/DOE-1.pdf}
\caption{\label{fig:DOE}Experimental design before disturbance. Communities
are implemented along a gradient of species richness (SR) and functional
dispersion (FDis) resulting in a broad range of aboveground biomass
(AGB). FDis was caluclated based on 4 functional traits (leaf mass per
area, wood specific gravity, maximum diameter, and maximum height).}
\end{figure}

\subsection{Ecosystem answer analysis}\label{ecosystem-answer-analysis}

\subsubsection{Resilience}\label{resilience}

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We decided to describe simiulated
tropical forests in three major functions: forest dynamic, forest
production, and forest diversity. Forest dynamic was represented by
aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\)
in \(m^2.ha^{-1}\)), total number of stem (\(N\)), number of stem above
\(10~cm\) diameter (\(N10\)), and number of stem above \(30~cm\)
diameter (\(N30\)). Forest production was represented by growth pimary
productivity (\(GPP\) in \(MgC.ha^{-1}\)), net pimary productivity
(\(NPP\) in \(MgC.ha^{-1}\)), tree autotrophic respiration in day
(\(Rday\) in \(MgC.ha^{-1}\)) and tree autotrophic respiration in night
(\(Rnight\) in \(MgC.ha^{-1}\)). Finally, forest diversity was
represented by \textbf{??}.

@@

\begin{equation}
  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
  \label{eq:Resilience}
\end{equation}

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
tropical forest ecosystems, the equilibrium used to calculate
\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
is dynamic. Consequently, to encompass undisturbed ecosystem variations
throught time, we simulated an undisturbed control ecosystem \(C\). And
the resilience of the system \(R(t)\) at the time \(t\) was defined as
the ratio of the ecosystem metric values in the disturbed simulation
\(X_T(t)\) over the the ecosystem metric values from the control
\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
when reaching the equilibrium value. Consequently we can calculate an
euclidean distance to equilibrium \(d_{eq}(t)\) as
\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\).

\subsubsection{Biodiversity effect}\label{biodiversity-effect}

Biodiversity is not only a facet of the experimental design and an
ecosystem output through forest diversity, but also interact on
ecosystem functioning and consequently on its answer to disturbance.
Biodiversity ecosystem functioning relation can be split in
complementarity and selection effect with \citet{Loreau2001}
partitioning:

@@

@@

\hypertarget{appendix-1-troll-model}{\section{Appendix 1: TROLL
model}\label{appendix-1-troll-model}}

@@

@@

\hypertarget{appendix-2-leaf-lifespan-model}{\section{Appendix 2: Leaf
lifespan model}\label{appendix-2-leaf-lifespan-model}}

TROLL model previous implementation encompass Reich's 1991 and 1997 and
Wright's 2004 allometries to estimate leaf lifespan with
\citep{Reich1991a, Reich1997, wright_worldwide_2004}. But we have shown
that Reich's allometries are underestimating leaf lifespan for low LMA
species. Moreover simulations estimated unrealistically low aboveground
biomass for low LMA species. We assumed Reich's allometries
underestimation of leaf lifespan for low LMA species being the source of
unrealistically low aboveground biomass inside TROLL simulations. We
decided to find a better allometry with \citet{wright_worldwide_2004}
GLOPNET dataset.

\subsection{Material and methods}\label{material-and-methods-1}

We gathered data from TRY database to evaluate and improve previous
models. We requested functional traits available in Bridge dataset and
leaf lifespan for tropical species present in GLOPNET dataset and used
in TROLL (see \ref{tab:A2traits}). We compiled GLOPNET, TRY, and DRYAD
databases. We kept dataset given by GLOPNET as origin dataset for
observations. It correspond to leaf lifespan and also to LMA and Nmass
in most of cases. We measured variable importance in functionnal traits
to explain leaf lifespan with out-of bag method applied on a random
forest.

\begin{table}

\caption{\label{tab:A2traits}Functional traits gathered with TRY.}
\centering
\begin{tabular}[t]{l|l|l|r}
\hline
Name & Trait & Unit & TRYcode\\
\hline
LL & Leaf lifespan (longevity) & month & 12\\
\hline
SLA & Leaf area per leaf dry mass (specific leaf area, SLA) & \$m\textasciicircum{}2.kg\textasciicircum{}\{-1\}\$ & 11\\
\hline
N & Leaf nitrogen (N) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 15\\
\hline
P & Leaf phosphorus (P) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 14\\
\hline
wsg & Stem dry mass per stem fresh volume (stem specific density) & \$mg.mm\textasciicircum{}\{-3\}\$ & 4\\
\hline
\end{tabular}
\end{table}

We tested different models and we kept the model with the best tradeoff
between model complexity, convergence, likelihood, and prediction
quality (see \href{LL2.html}{leaf lifespan model} document). We found
following model, named \textbf{M1}:
\[M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,\]
\[s=1,...,S_{=4}~, ~~j=1,...,n_s\]
\[{\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}\]

Leaf lifespan (\(LL\)) is following a lognormal law with location
\({\mu_s}_j\) and scale \(\sigma\). \(s\) represents the dataset used to
do the fit between 1 and \(S=4\), it encompass environmental and
protocol variations. \(j\) represents the observation for a given site
\(s\). Location \({\mu_s}_j\) is the exponential from the difference
between LMA and Nmass with parameters \({\beta_0}_s\), \({\beta_1}_s\),
and \({\beta_2}_s\). Each \({\beta_i}_s\) is following a normal law
located on \(\beta_i\) with scale \(\sigma_i\). Finally all \(beta_i\),
\(\sigma\), and \(\sigma_i\) are assumed without presemption following
gamma laws of parameters \(10^-2\), \(10^-2\).

We decided to add to model \textbf{M1}, wood specific gravity as a new
predictor through a parameter \(\beta_3\) in a new model named
\textbf{M10}. We removed dataset random effect on \(\beta_3\) due to the
fact that all wood specific gravities were provided by DRYAD database.
Due to a correlation between \(\beta_0\) and \(\beta_3\) in \textbf{M10}
(see figure 5), we tried also a model \textbf{M11} without intercept
\(\beta_0\). We then sampled all models with this composite dataset, and
obtained following results :

\subsection{Results}\label{results-1}

\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tab:OOB}Variable importance calculated with out-of the bag
method applied on a random forest. Leaf lifespan (LL) is taken in
GLOPNET database from \citet{wright_worldwide_2004}. Leaf mass per area
(LMA), and leaf nitrogen content (Nmass) are taken both in TRY
(\url{https://www.try-db.org}) and GLOPNET databases. Wood specific
gravity (wsg) is taken both in TRY and DRYAD databases.}\tabularnewline
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endfirsthead
\toprule
& \%IncMSE & IncNodePurity\tabularnewline
\midrule
\endhead
LMA & 99.69390 & 2028.079\tabularnewline
Nm & 159.03360 & 2666.670\tabularnewline
wsg & 50.97284 & 1475.023\tabularnewline
\bottomrule
\end{longtable}

Table 3: Model fit results for composite dataset. \emph{Data are taken
from GLOPNET, TRY, and DRYAD datasets (see previous description). ML
stand for maximum likelihood and RMSEP for root mean square error of
prediction}

All model have correctly converged.

Model \textbf{M10} have shown a correlation between parameters
\(\beta_0\) and \(\beta_3\), responsible for the creation of model
\textbf{M11} (see previous paragraphs). Model \textbf{M11} only show a
slight link between \(\beta_1\) and \(\beta_3\) which can be considered
as non significant.

We can denote that retained vector of parameters with maximum likelihood
for model \textbf{M11} is close to 95\% confidence limit.

We finally tested the new allometry obtained with the new model
\textbf{M} on TROLL data (figure 7) and with TROLL simulations. Minimum
leaf lifespan is now above 3 months for the new model \textbf{M} with
TROLL species dataset. Moreover lowest leaf lifespan species are not the
one with the smallest LMA with new model. Lastly, new model \textbf{M}
variation of leaf lifespan is almost spanning all previous allometries
prediciton on LMA axis.

Simulations are validating that this new allometry resolve the issue of
unrealistically low aboveground biomass for low LMA species due to an
early edath of individuals inside simulations. For instance with this
allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
aboveground biomass above X tonnes of C/ha and realistic diameter and
age distribution inside the final population.

\hypertarget{appendix-3-rotten-tree-model}{\section{Appendix 3: Rotten
tree model}\label{appendix-3-rotten-tree-model}}

\hypertarget{appendix-4-gap-damages-model}{\section{Appendix 4: Gap
damages model}\label{appendix-4-gap-damages-model}}
